---
title: "Deepak Mongia - Data 607 Week 9 Assignment"
author: "Deepak Mongia"
date: "October 27, 2018"
output: 
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
---

## Assignment - Web APIs

The New York Times web site provides a rich set of APIs, as described here: http://developer.nytimes.com/docs
You'll need to start by signing up for an API key.
Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it to an R dataframe.


Step-1: Loading the required packages
```{r}
library(httr)
library(jsonlite)
library(dplyr)
library(kableExtra)
```

We have already requested the API key to get / read the data from the NY Times API. We are using the books_api from the NY Times page : http://developer.nytimes.com/

```{r}
nyt_books_url <- "https://api.nytimes.com/svc/books/v3/lists/best-sellers/history.json"

nyt_books <- GET(nyt_books_url, query = list(api_key = "a01880c89030475c908b46f8072e2a02"))

nyt_books$status_code

text_content <- content(nyt_books, as = "text")

json_content_df <- text_content %>% fromJSON() %>% as.data.frame()

dim(json_content_df)

json_content_df %>% kable() %>% kable_styling()

# Removing the first 3 columns as those are irrelevant here, and not needed
json_content_df <- json_content_df[, -(1:3)]

json_content_df %>% kable() %>% kable_styling()

```

