---
title: "Deepak Mongia Data607 - Week 5 Assignment"
author: "Deepak Mongia"
date: "September 30, 2018"
output: html_document
---

Assignment - Tidying and Transforming Data 

1. Loading the required packages

```{r, echo=FALSE}
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)
library(ggplot2)
```

2. Reading the file into a data.frame from the Raw Github link
```{r}
getURL <- "https://raw.githubusercontent.com/deepakmongia/Fall2018/master/Arrival_Delays.csv"
arrival.delays.df <- read.csv(getURL, header = TRUE, sep = ",")
```

3. 
```{r}
colnames(arrival.delays.df)[colnames(arrival.delays.df) == "X"] <- "airline"
colnames(arrival.delays.df)[colnames(arrival.delays.df) == "X.1"] <- "status"
arrival.delays.df

# Removing the blank lines
arrival.delays.df <- arrival.delays.df %>% filter(status == "on time" | status == "delayed")
arrival.delays.df
```

4. Replacing the blank values in the column - airlines with their flight name
```{r, echo=FALSE}
airline_names_vector <- arrival.delays.df[!is.na(str_extract(arrival.delays.df$airline, "\\w+")), ]$airline
arrival.delays.df[is.na(str_extract(arrival.delays.df$airline, "\\w+")), ]$airline <- airline_names_vector
arrival.delays.df
```

5. 
```{r}
arrival.delays.df.molten <- arrival.delays.df %>% gather("destination", "frequency", 3:7)
arrival.delays.df.molten
arrival.delays.df.tidy <- arrival.delays.df.molten[, c(1,3,2,4)]
arrival.delays.df.tidy
```

Now if we see the above data set, it is a tidy one, as each column is a variable, and each row is an observation.

We will be doing some analysis now on this tidy data set.

```{r, echo=FALSE}
(airline_citywise_stats <- group_by(arrival.delays.df.tidy, airline, destination, status))
airline_citywise_breakup <- summarize(airline_citywise_stats, count=sum(frequency))
airline_citywise_breakup

airline_citywise_breakup_Alaska <- filter(airline_citywise_breakup, airline == "ALASKA")
airline_citywise_breakup_AMWest <- filter(airline_citywise_breakup, airline == "AM WEST")

airline_citywise_breakup_Alaska
airline_citywise_breakup_AMWest

```

Counts break-up of the 2 airlines for city wise on-time and delayed flight counts
```{r, echo=FALSE}
ggplot(airline_citywise_breakup_Alaska, aes(x=destination, y=count, fill=status)) + geom_bar(stat = "identity") + labs(title="Alaska Airlines destination wise break-up of counts")
ggplot(airline_citywise_breakup_AMWest, aes(x=destination, y=count, fill=status)) + geom_bar(stat = "identity") + labs(title="AM West Airlines destination wise break-up of counts")
```

Percentage break-ups of the 2 airlines, ratio wise break-ups for each city 
```{r, echo=FALSE}
#For Alaska Airlines
airline_citywise_breakup_Alaska_percent <- ddply(airline_citywise_breakup_Alaska, "destination", transform, percent_count=count/sum(count) * 100)

ggplot(airline_citywise_breakup_Alaska_percent, aes(x=destination, y=percent_count, fill=status)) + geom_bar(stat = "identity") + labs(title="Alaska Airlines destination wise break-up of counts - by ratio")

#For AM West Airlines
airline_citywise_breakup_AMWest_percent <- ddply(airline_citywise_breakup_AMWest, "destination", transform, percent_count=count/sum(count) * 100)

ggplot(airline_citywise_breakup_AMWest_percent, aes(x=destination, y=percent_count, fill=status)) + geom_bar(stat = "identity") + labs(title="AM West Airlines destination wise break-up of counts - by ratio")

```

Conclusion:

1. From the last 2 graphs - ratio break-ups for each city signifies that for both the airlines, each city has almost similar trend for the 2 airlies when the ratio of the delayed to the on-time arrivals are considered. 

2. However, the it is also quite visible from the graphs that AM West Airlines also has more chance of a delay as compared to the Alaska Airlines for any of the 5 given cities for which the readings have been considered for this analysis.
