---
title: "Deepak Mongia - Data 607 - Project-2"
author: "Deepak Mongia"
date: "October 7, 2018"
output: html_document
---

This is the Project-2 in DATA607 course for CUNY SPS -Masters in Data Science program.
As a part of this project, 3 wide or messy data sets are being read, cleaned up, and then used for individual analyses.

Loading the required packages.
```{r setup, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(stringr)
```

Data set # 1 - UNICEF data for number of deaths of children age - 5-14 years in multiple areas or geographies
We are now reading the first data set.
```{r}
getURL <- "https://raw.githubusercontent.com/deepakmongia/Fall2018/master/Number_of_deaths_estimates-childresn-age-5-14.csv"

Children_death_estimates_df <- read.csv(getURL, header = TRUE, sep = ",", skip = 1)

head(Children_death_estimates_df)
```

Now we are melting the data set, that means all the columns which are data are being converted to separate columns.
```{r}
Children_death_estimates_df_molten <- Children_death_estimates_df %>% gather("year", "Number.of.deaths", 3:7)

head(Children_death_estimates_df_molten)
```

```{r}
Children_death_estimates_df_molten$year <- substr(Children_death_estimates_df_molten$year, 2, 5)
Children_death_estimates_df_molten$Number.of.deaths <- gsub(",", "", Children_death_estimates_df_molten$Number.of.deaths)

head(Children_death_estimates_df_molten)
```

```{r}
Children_death_estimates_df_molten_median <- Children_death_estimates_df_molten %>% filter(Children_death_estimates_df_molten$Uncertainty.bounds. == "Median", Children_death_estimates_df_molten$Region.Name != "World")

head(Children_death_estimates_df_molten_median)
```

```{r}
Children_death_estimates_df_molten_median <- Children_death_estimates_df_molten_median %>% arrange(Region.Name, year)
```




DATASET-2

The secod data set is for the country wise data for Balance of payment as a percent of that country's GDP

Step-1 - To read the data set from Github location into R
```{r}
getURL2 <- "https://raw.githubusercontent.com/deepakmongia/Fall2018/master/Selected_Indicators_Percent_of_GDP.csv"
Selected_Indicators_Percent_of_GDP_raw <- read.csv(getURL2, header = TRUE, sep = ",", skip = 3)
```

Performing the steps to
1) rename the first colum  to Country.names
2) Removing the junk columns from the end
3) Replcaing the unknown data given by 3 dots "..." by NA
```{r}
head(Selected_Indicators_Percent_of_GDP_raw)
colnames(Selected_Indicators_Percent_of_GDP_raw)[colnames(Selected_Indicators_Percent_of_GDP_raw) == "X"] = "Country.name"
names(Selected_Indicators_Percent_of_GDP_raw)

Selected_Indicators_Percent_of_GDP_raw <- Selected_Indicators_Percent_of_GDP_raw %>% select(1:9)
Selected_Indicators_Percent_of_GDP_raw[Selected_Indicators_Percent_of_GDP_raw == "..."] <- NA
```

Pull only the continents / categories and countries with non-blank values into a new filtered data set
```{r}
Selected_Indicators_Percent_of_GDP_filtered <-  Selected_Indicators_Percent_of_GDP_raw %>% filter((substr(Country.name,1,8) == "Advanced" | substr(Country.name,1,4) == "Euro"
                                                   | substr(Country.name,1,8) == "Emerging" | substr(Country.name,1,3) == "CIS"
                                                   | substr(Country.name,1,6) == "Middle" | substr(Country.name,1,11) == "Sub-Saharan"
                                                   | substr(Country.name,1,5) == "CEMAC" | substr(Country.name,1,5) == "WAEMU" 
                                                   | substr(Country.name,1,7) == "Western" | substr(Country.name,1,4) == "ECCU")
                                                  | !(is.na(X2009) & is.na(X2010) & is.na(X2011) & is.na(X2012) & is.na(X2013) & is.na(X2014) & is.na(X2015) & is.na(X2016)))


```

Creating a new column called category, this column will be filled with the category under which any country lies. For example - Advanced economies
```{r}
Selected_Indicators_Percent_of_GDP_filtered$Category <- Selected_Indicators_Percent_of_GDP_filtered$Country.name

end_of_df <- "FALSE"
i <- 1

while (end_of_df == "FALSE") {
  if (Selected_Indicators_Percent_of_GDP_filtered$Country.name[i] %>% is.na())
  {
    end_of_df <- "TRUE"
  }
  else
    #  sprintf("not end of file for record # %s", i)
  {
    if (Selected_Indicators_Percent_of_GDP_filtered$X2009[i] %>% is.na() &
        Selected_Indicators_Percent_of_GDP_filtered$X2010[i] %>% is.na() &
        Selected_Indicators_Percent_of_GDP_filtered$X2011[i] %>% is.na() &
        Selected_Indicators_Percent_of_GDP_filtered$X2012[i] %>% is.na() &
        Selected_Indicators_Percent_of_GDP_filtered$X2013[i] %>% is.na() &
        Selected_Indicators_Percent_of_GDP_filtered$X2014[i] %>% is.na() &
        Selected_Indicators_Percent_of_GDP_filtered$X2015[i] %>% is.na() &
        Selected_Indicators_Percent_of_GDP_filtered$X2016[i] %>% is.na())
    {
      save.country.category <- Selected_Indicators_Percent_of_GDP_filtered$Country.name[i]
    }
    else
    {
      Selected_Indicators_Percent_of_GDP_filtered$Category[i] <- save.country.category
    }
    i <- i + 1 
  }
}

```


Only saving the country rows and removing the category rows, as we now have grabbed the category values in a new column for all the countries
```{r}
Selected_Indicators_Percent_of_GDP_filtered <- Selected_Indicators_Percent_of_GDP_filtered %>% filter(!(Selected_Indicators_Percent_of_GDP_filtered$X2009 %>% is.na() &
                                                                                                          Selected_Indicators_Percent_of_GDP_filtered$X2010 %>% is.na() &
                                                                                                          Selected_Indicators_Percent_of_GDP_filtered$X2011 %>% is.na() &
                                                                                                          Selected_Indicators_Percent_of_GDP_filtered$X2012 %>% is.na() &
                                                                                                          Selected_Indicators_Percent_of_GDP_filtered$X2013 %>% is.na() &
                                                                                                          Selected_Indicators_Percent_of_GDP_filtered$X2014 %>% is.na() &
                                                                                                          Selected_Indicators_Percent_of_GDP_filtered$X2015 %>% is.na() &
                                                                                                          Selected_Indicators_Percent_of_GDP_filtered$X2016 %>% is.na()))


head(Selected_Indicators_Percent_of_GDP_filtered)
```

Bringing the category on the left before the data set is melted. And then melting the data set.
```{r}
Selected_Indicators_Percent_of_GDP_filtered <- Selected_Indicators_Percent_of_GDP_filtered[, c(1,10,2:9)]

Selected_Indicators_Percent_of_GDP_long <- Selected_Indicators_Percent_of_GDP_filtered %>% gather("year", "BOP_Percent_of_GDP", 3:10)

head(Selected_Indicators_Percent_of_GDP_long)
```

```{r}
Selected_Indicators_Percent_of_GDP_long$year <- substr(Selected_Indicators_Percent_of_GDP_long$year, 2,5)

Selected_Indicators_Percent_of_GDP_long$BOP_Percent_of_GDP <- as.numeric(Selected_Indicators_Percent_of_GDP_long$BOP_Percent_of_GDP)

BOP_by_GDP_category_year_summarized <- Selected_Indicators_Percent_of_GDP_long %>% group_by(Category, year) %>% summarize(avg_percent_BOP=round(mean(BOP_Percent_of_GDP, na.rm = TRUE),2))

ggplot(BOP_by_GDP_category_year_summarized, aes(x=year, y=avg_percent_BOP, fill=Category)) + geom_bar(position = "dodge", stat = "identity")
```

