---
title: "Data 607 Project-3 - freeformResponses Cleaning"
author: "Deepak Mongia"
date: "October 16, 2018"
output: html_document
---
  
```{r setup, include=FALSE}
library(dplyr)
library(stringr)
```

```{r}
getURL <- "https://raw.githubusercontent.com/deepakmongia/Fall2018/master/Data-607-Project3/freeformResponses.csv"
free.form.response.df1 <- read.csv(getURL, header = TRUE, sep = ",")

dim(free.form.response.df1)

head(free.form.response.df1, 2)
```

Removing the columns which have all NA values
```{r}
free.form.response.df2 <- free.form.response.df1[, apply(free.form.response.df1, 2, function(x){any(!is.na(x))})]
```

Checking the counts of valid data elements under all the columns in the new dataframe
```{r}
sprintf("Total number of rows in the data.frame is: %s", nrow(free.form.response.df2))
apply(free.form.response.df2, 2, function(x){sum(!(x == ""), na.rm = TRUE)})
```

```{r}
KaggleMotivationFreeForm_vect <- free.form.response.df2$KaggleMotivationFreeForm[free.form.response.df2$KaggleMotivationFreeForm != ""]
length(KaggleMotivationFreeForm_vect)

CurrentJobTitleFreeForm_vect <- free.form.response.df2$CurrentJobTitleFreeForm[free.form.response.df2$CurrentJobTitleFreeForm != ""]
length(CurrentJobTitleFreeForm_vect)

any(is.na(free.form.response.df2$CurrentJobTitleFreeForm))

MLToolNextYearFreeForm_vect <- free.form.response.df2$MLToolNextYearFreeForm[free.form.response.df2$MLToolNextYearFreeForm != ""]
length(MLToolNextYearFreeForm_vect)
MLToolNextYearFreeForm_vect

MLMethodNextYearFreeForm_vect <- free.form.response.df2$MLMethodNextYearFreeForm[free.form.response.df2$MLMethodNextYearFreeForm != ""]
length(MLMethodNextYearFreeForm_vect)

LanguageRecommendationFreeForm_vect <- free.form.response.df2$LanguageRecommendationFreeForm[free.form.response.df2$LanguageRecommendationFreeForm != ""]
length(LanguageRecommendationFreeForm_vect)

JobSkillImportanceOtherSelect1FreeForm_vect <- free.form.response.df2$JobSkillImportanceOtherSelect1FreeForm[free.form.response.df2$JobSkillImportanceOtherSelect1FreeForm != ""]
length(JobSkillImportanceOtherSelect1FreeForm_vect)

JobSkillImportanceOtherSelect2FreeForm_vect <- free.form.response.df2$JobSkillImportanceOtherSelect2FreeForm[free.form.response.df2$JobSkillImportanceOtherSelect2FreeForm != ""]

JobSkillImportanceOtherSelect3FreeForm_vect <- free.form.response.df2$JobSkillImportanceOtherSelect3FreeForm[free.form.response.df2$JobSkillImportanceOtherSelect3FreeForm != ""]

PastJobTitlesFreeForm_vect <- free.form.response.df2$PastJobTitlesFreeForm[free.form.response.df2$PastJobTitlesFreeForm != ""]
length(PastJobTitlesFreeForm_vect)

WorkHardwareFreeForm_vect <- free.form.response.df2$WorkHardwareFreeForm[free.form.response.df2$WorkHardwareFreeForm != ""]

SalaryChangeFreeForm_vect <- free.form.response.df2$SalaryChangeFreeForm[free.form.response.df2$SalaryChangeFreeForm != ""]

TimeOtherSelectFreeForm_vect <- free.form.response.df2$TimeOtherSelectFreeForm[free.form.response.df2$TimeOtherSelectFreeForm != ""]
length(TimeOtherSelectFreeForm_vect)


## counts for programming being devoted time and its data

sum(!is.na(str_extract(TimeOtherSelectFreeForm_vect, "(\\w*rogramming)|(\\w*oftware)|(\\w*lgorithm)|(\\w*coding)|(\\w*Coding)|(\\w*development)")))

TimeOtherSelectFreeForm_vect[!is.na(str_extract(TimeOtherSelectFreeForm_vect, "(\\w*rogramming)|(\\w*oftware)|(\\w*lgorithm)|(\\w*Coding)|(\\w*development)"))]


## counts for management being used based on time and its values
sum(!is.na(str_extract(TimeOtherSelectFreeForm_vect, "\\w*anage")))
TimeOtherSelectFreeForm_vect[!is.na(str_extract(TimeOtherSelectFreeForm_vect, "\\w*anage"))]


```


Creating a data.frame for just the non blank data from the relevant columns
```{r}
FreeForm_List <- list(KaggleMotivationFreeForm = KaggleMotivationFreeForm_vect, CurrentJobTitleFreeForm = CurrentJobTitleFreeForm_vect, MLToolNextYearFreeForm = MLToolNextYearFreeForm_vect, MLMethodNextYearFreeForm = MLMethodNextYearFreeForm_vect, LanguageRecommendationFreeForm = LanguageRecommendationFreeForm_vect, JobSkillImportanceOtherSelect1FreeForm = JobSkillImportanceOtherSelect1FreeForm_vect, JobSkillImportanceOtherSelect2FreeForm = JobSkillImportanceOtherSelect2FreeForm_vect, JobSkillImportanceOtherSelect3FreeForm = JobSkillImportanceOtherSelect3FreeForm_vect, PastJobTitlesFreeForm = PastJobTitlesFreeForm_vect, WorkHardwareFreeForm = WorkHardwareFreeForm_vect, SalaryChangeFreeForm = SalaryChangeFreeForm_vect, TimeOtherSelectFreeForm = TimeOtherSelectFreeForm_vect)


FreeForm_df <- FreeForm_List %>% lapply(`length<-`, max(lengths(FreeForm_List))) %>% as.data.frame()

head(FreeForm_df)


```


Now checking the job titles of the people who selected the "other" option from the job title form, and hence fall under the free form - CurrentJobTitleFreeForm:
```{r}
CurrentJobTitleFreeForm_vect <- tolower(CurrentJobTitleFreeForm_vect)
CurrentJobTitleFreeForm_vect %>% as.factor() %>% levels()

### Data related jobs
sum(!is.na(str_extract(CurrentJobTitleFreeForm_vect, "(\\w*business intelligence)|(\\w*bi)|(\\w*data)|(\\w*analyst)|(\\w*big data)|(\\w*dba)")))
CurrentJobTitleFreeForm_vect[!is.na(str_extract(CurrentJobTitleFreeForm_vect, "(\\w*business intelligence)|(\\w*bi)|(\\w*data)|(\\w*analyst)|(\\w*big data)|(\\w*dba)"))]


### software related jobs
sum(!is.na(str_extract(CurrentJobTitleFreeForm_vect, "(\\w*application)|(\\w*computer)|(\\w*software)|(^it)")))
CurrentJobTitleFreeForm_vect[!is.na(str_extract(CurrentJobTitleFreeForm_vect, "(\\w*application)|(\\w*computer)|(\\w*software)|(^it)"))]

### leadership using data science
sum(!is.na(str_extract(CurrentJobTitleFreeForm_vect, "(\\w*director)|(\\w*cto)|(\\w*ceo)|(\\w*cfo)|(\\w*chief)|(\\w*vp)|(\\w*head of)")))
CurrentJobTitleFreeForm_vect[!is.na(str_extract(CurrentJobTitleFreeForm_vect, "(\\w*director)|(\\w*cto)|(\\w*ceo)|(\\w*cfo)|(\\w*chief)|(\\w*vp)|(\\w*head of)"))]

### Teching / Education
sum(!is.na(str_extract(CurrentJobTitleFreeForm_vect, "(\\w*professor)|(\\w*teach)|(\\w*educat)|(\\w*faculty)")))
CurrentJobTitleFreeForm_vect[!is.na(str_extract(CurrentJobTitleFreeForm_vect, "(\\w*professor)|(\\w*teach)|(\\w*educat)|(\\w*faculty)"))]

### Student
sum(!is.na(str_extract(CurrentJobTitleFreeForm_vect, "(\\w*student)")))
CurrentJobTitleFreeForm_vect[!is.na(str_extract(CurrentJobTitleFreeForm_vect, "(\\w*student)"))]

### Finance related
sum(!is.na(str_extract(CurrentJobTitleFreeForm_vect, "(\\w*financ)")))
CurrentJobTitleFreeForm_vect[!is.na(str_extract(CurrentJobTitleFreeForm_vect, "(\\w*financ)"))]

### Quant / maths related
sum(!is.na(str_extract(CurrentJobTitleFreeForm_vect, "(\\w*quant)|(\\w*math)")))
CurrentJobTitleFreeForm_vect[!is.na(str_extract(CurrentJobTitleFreeForm_vect, "(\\w*quant)|(\\w*math)"))]

```

So, that clearly shows that the other job title selected in the survey falls mainly under below 7 categories:

1) Data Related jobs (like data engineer, data analytics, etc.)

2) Software related jobs

3) Executive or Top leadership positions in any organization / Institution (like ceo, cfo, head of engineering, etc.)

4) Teaching / Education

5) Student

6) Finance related job profiles

7) Quantitative or Maths related job profiles
