---
title: "Data 606 Homework-5"
author: "Deepak Mongia"
date: "October 29, 2018"
output: html_document
---

5.6 Working backwards, Part II. A 90% confidence interval for a population mean is (65,
77). The population distribution is approximately normal and the population standard deviation
is unknown. This confidence interval is based on a simple random sample of 25 observations.
Calculate the sample mean, the margin of error, and the sample standard deviation.

Answer:
As we are calculating 90% confidence level here, that means 5% on each side. df = n-1, hence df = 24 in this case
Calculating the t-value:
```{r}
t5.6 <- abs(qt(p=0.05, df=24))
t5.6
```

Calculating the standard error:
```{r}
SE_5.6 <- (77-65)/(2 * t5.6)
SE_5.6
```

Calculating the standard deviation
```{r}
sd_5.6 <- SE_5.6 * sqrt(25)
sd_5.6
```

Calculating the Margin of error:
```{r}
ME_5.6 <- t5.6 * SE_5.6
ME_5.6
```


5.14 SAT scores. SAT scores of students at an Ivy League college are distributed with a
standard deviation of 250 points. Two statistics students, Raina and Luke, want to estimate the
average SAT score of students at this college as part of a class project. They want their margin of
error to be no more than 25 points.
(a) Raina wants to use a 90% confidence interval. How large a sample should she collect?

Answer: 
Calculating z-score for a normal distribution for 95% CI
```{r}
sd_5.14 <- 250
Z_5.14_Raina <- qnorm(0.95, mean = 0, sd = 1)
```

Deriving the formula for this question.
now, Margin Error (ME) < 25
Z * SE < 25

Z * (sd/sqrt(n))  < 25

n > (Z * sd / 25)^2
Using this in R:
```{r}
n_5.14_Raina <- (Z_5.14_Raina * sd_5.14 / 25)^2
n_5.14_Raina
```

Hence Raina should collect at least 271

(b) Luke wants to use a 99% confidence interval. Without calculating the actual sample size,
determine whether his sample should be larger or smaller than Raina's, and explain your
reasoning.

Answer: 
Calculating z-score for a normal distribution for 99% CI
```{r}
Z_5.14_Luke <- qnorm(0.995, mean = 0, sd = 1)
Z_5.14_Luke
```

Bigger the Z-score, bigger the sample. As Luke's Z-score is bigger, Luke's sample will have to be larger.

(c) Calculate the minimum required sample size for Luke.

Answer:

```{r}
#Using the same formula for Luke - using Z-score for Luke
n_5.14_Luke <- (Z_5.14_Luke * sd_5.14 / 25)^2
n_5.14_Luke
```