---
title: "Data 606 Chapter-7 Homework"
author: "Deepak Mongia"
date: "November 25, 2018"
output: html_document
---

## 7.24 Nutrition at Starbucks, Part I. 
The scatterplot below shows the relationship between
the number of calories and amount of carbohydrates (in grams) Starbucks food menu items contain.
21 Since Starbucks only lists the number of calories on the display items, we are interested in
predicting the amount of carbs a menu item has based on its calorie content.

(a) Describe the relationship between number of calories and amount of carbohydrates (in grams)
that Starbucks food menu items contain.

Answer:
From the graph shown, it is very clear that there is a strong positive linear relationship between the 2 variables - calories and carbs in Starbucks food items.

(b) In this scenario, what are the explanatory and response variables?

Answer:
In this case, Calories is the explanatory variable and carbs is the response variable

(c) Why might we want to fit a regression line to these data?

Answer: As we see that there is a strong positive relationship between the 2 variables, we would like to draw a regression line between these 2 variables. It will help us understand and predict how the presence of calories determines the quantity of carbs in a food item.

(d) Do these data meet the conditions required for fitting a least squares line?

Answer: For the least suqared lines fitting, we will check how the 4 conditions for the least squared line fair for this scenario:
1) Linear: The relationship is linear as we can see from the scatterplot between the 2 variables.
2) Nearly normal residuals check: The residuals plot clearly shows that the residuals is not close to normal.
3) Constant variability: It is clear from the residual plot that there are outliers on wither side of the residual plots.
4) Indepdendent observations: These are all independent observations as these are for different products.

Even though 1 and 4 above meet the least square line criteria, but 2 and 3 are quite off the mark to say that this scenario meets the least squared line criteria.



## 7.26 Body measurements, Part III. 
Exercise 7.15 introduces data on shoulder girth and
height of a group of individuals. The mean shoulder girth is 107.20 cm with a standard deviation
of 10.37 cm. The mean height is 171.14 cm with a standard deviation of 9.41 cm. The correlation
between height and shoulder girth is 0.67.

(a) Write the equation of the regression line for predicting height.

Answer:
Here shoulder girth is the explanatory variable, x
and height is the response variable, y

x_bar = 107.2
s_x = 10.37

y_bar = 171.14
s_y = 9.41

R = 0.67

slope, b1 = (s_y / s_x) * R
```{r}
x_bar_7.26 <- 107.2
s_x_7.26 <- 10.37

y_bar_7.26 <- 171.14
s_y_7.26 <- 9.41

R_7.26 <- 0.67


#slope
b1_7.26 = (s_y_7.26 / s_x_7.26) * R_7.26
b1_7.26

```

Equation would be:
y - y_bar_7.26 = b1_7.26 * (x - x_bar_7.26)
y - 171.14 = 0.608 * (x - 107.2)
y = 105.96 + 0.608 x

(b) Interpret the slope and the intercept in this context.

Answer : Intercept value = 105.96
That signifies that when x (shoulder girth) = 0, y (height) = 105.96
That does not sound to be a valid scenario, Hence we can say that suggests the height of the line

Slope = 0.608 means that for every 1 cm of shoulder girth, the height increases by 1 cm.

(c) Calculate R2 of the regression line for predicting height from shoulder girth, and interpret it
in the context of the application.

Answer:

```{r}
(R_7.26)^2
```

That means around 44.9% of the variability in height is measured by the shoulder girth.
More the value of R^2 or R, more linear and clean the relationship.

(d) A randomly selected student from your class has a shoulder girth of 100 cm. Predict the height
of this student using the model.

Answer:
```{r}
y_7.26_d <- 105.96 + 0.608 * 100
y_7.26_d
```

(e) The student from part (d) is 160 cm tall. Calculate the residual, and explain what this residual
means.

Answer:
```{r}
y_7.26_oberved <- 160
residual_7.26 <- y_7.26_oberved - y_7.26_d
residual_7.26
```

Residual is -6.76, that means the observed value is less than the model calculated value, or the model over-predicted the value.

(f) A one year old has a shoulder girth of 56 cm. Would it be appropriate to use this linear model
to predict the height of this child?

Answer:
No, it would not be appropriate to use the model here, as we would be extrapolating as the shoulder girth of 56 cm is outside the observed range.



## 7.30 Cats, Part I. 
The following regression output is for predicting the heart weight (in g) of cats
from their body weight (in kg). The coefficients are estimated using a dataset of 144 domestic cats.

(a) Write out the linear model.

Answer:
y = -0.357 + 4.034 x

(b) Interpret the intercept.

Answer:
When x (body weight) = 0, y (heart weight) = -0.357, that is not a valid scenario.
Hence in this scenario, the intercept determines the height of the linear regression line.

(c) Interpret the slope.

Answer: 
For every 1 kg increase of body weight, the heart weight increases by 4.034

(d) Interpret R^2.

Answer: 
Body weights account for 64.66% of variability in heart weight.

(e) Calculate the correlation coefficient.

Answer:
```{r}
R2_7.30 <- 0.6466
R_7.30 <- sqrt(R2_7.30)
R_7.30
```

Correlation coefficient R = 0.804


