---
title: "Data 606 Project Proposal"
author: "Deepak Mongia"
date: "October 27, 2018"
output: html_document
---
### Energy efficiency Data Set

For this project for Data 606, we are going to work on the Energy Efficient data set which is present at this web link: http://archive.ics.uci.edu/ml/datasets/Energy+efficiency#

### Data Preparation

Loading the data set into R, so that it can be used
```{r setup}
library(readxl)
library(ggplot2)
library(gridExtra)

energy.efficiency.df1 <- read_xlsx("ENB2012_data.xlsx", sheet = 1)

head(energy.efficiency.df1)

names(energy.efficiency.df1) <- c("relative.compactness", "surface.area", "wall.area", "roof.area", "overall.height", "orientation", "glazing.area", "glazing.area.distribution", "heating.load", "cooling.load")

summary(energy.efficiency.df1)

energy.efficiency.df1$orientation <- as.factor(energy.efficiency.df1$orientation)
levels(energy.efficiency.df1$orientation)[levels(energy.efficiency.df1$orientation) == "2"] <- "north"
levels(energy.efficiency.df1$orientation)[levels(energy.efficiency.df1$orientation) == "3"] <- "east"
levels(energy.efficiency.df1$orientation)[levels(energy.efficiency.df1$orientation) == "4"] <- "south"
levels(energy.efficiency.df1$orientation)[levels(energy.efficiency.df1$orientation) == "5"] <- "west"

head(energy.efficiency.df1)

```

Attribute Details:

Relative Compactness        -     Ratio

Surface Area                -     sq. meters

Wall Area                   -     sq. meters

Roof Area                   -     sq. meters

Overall Height              -     meters

Orientation                 -     2:North, 3:East, 4:South, 5:West

Glazing area (ratio)        -     0.00, 0.10, 0.25, 0.40

Glazing area distribution   -     1:Uniform, 2:North, 3:East, 4:South, 5:West

Heating Load                -     kWh/sq. meters

Cooling Load                -     kWh/sq. meters


### Research question

**You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.**

1. Is the relative compactness of a building predictive of heating load and cooling load of a building ?

2. Is the area (wall area / roof area / surface area) of a building predictive of the heating or cooling load ?

3. Does the building orientation suggest anything about the heating or cooling efficiency of a building ?



### Cases

**What are the cases, and how many are there?**

- Cases are all these variable values for each building which is a part of the sample

- There are total of 768 cases in this data set

**Describe the method of data collection.**

The dataset was created by Angeliki Xifara (angxifara '@' gmail.com, Civil/Structural Engineer) and was processed by Athanasios Tsanas (tsanasthanasis '@' gmail.com, Oxford Centre for Industrial and Applied Mathematics, University of Oxford, UK).


### Type of study 

**What type of study is this (observational/experiment)?**

This is an observational study as the data has been collected for the buildings which are there already. This is not an experiment.


### Data Source 

**If you collected the data, state self-collected. If not, provide a citation/link.**

The data is collected from: http://archive.ics.uci.edu/ml/datasets/Energy+efficiency#

### Dependent Variable

**What is the response variable? Is it quantitative or qualitative?**

The response variables are:
heating.load
cooling.load


### Independent Variable

**You should have two independent variables, one quantitative and one qualitative.**

In this case, we have the below independent variables:
relative.compactness      -   numerical
surface.area              -   numerical
wall.area                 -   numerical
roof.area                 -   numerical
overall.height            -   numerical
orientation               -   categorical
glazing.area              -   numerical
glazing.area.distribution -   categorical


### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

Efficiencies against relative compactness
```{r}
a <- ggplot(energy.efficiency.df1, aes(x=energy.efficiency.df1$relative.compactness, y=energy.efficiency.df1$heating.load)) +
  geom_point()

b <- ggplot(energy.efficiency.df1, aes(x=energy.efficiency.df1$relative.compactness, y=energy.efficiency.df1$cooling.load)) +
  geom_point()

grid.arrange(a, b, nrow = 1)
```

Efficiencies against surface area
```{r}
a <- ggplot(energy.efficiency.df1, aes(x=energy.efficiency.df1$surface.area, y=energy.efficiency.df1$heating.load)) +
  geom_point()

b <- ggplot(energy.efficiency.df1, aes(x=energy.efficiency.df1$surface.area, y=energy.efficiency.df1$cooling.load)) +
  geom_point()

grid.arrange(a, b, nrow = 1)

```

Relationships between the 2 efficiency values
```{r}
ggplot(energy.efficiency.df1, aes(x=energy.efficiency.df1$heating.load, y=energy.efficiency.df1$cooling.load)) +
  geom_point() + stat_smooth(method = lm)
```

