---
title: 'Inference for numerical data'
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---
```{r}
library(ggplot2)
library(dplyr)
```


## North Carolina births

In 2004, the state of North Carolina released a large data set containing 
information on births recorded in this state. This data set is useful to 
researchers studying the relation between habits and practices of expectant 
mothers and the birth of their children. We will work with a random sample of 
observations from this data set.

## Exploratory analysis

Load the `nc` data set into our workspace.

```{r load-data, eval=TRUE}
load("more/nc.RData")
```

We have observations on 13 different variables, some categorical and some 
numerical. The meaning of each variable is as follows.

variable         | description
---------------- | -----------
`fage`           | father's age in years.
`mage`           | mother's age in years.
`mature`         | maturity status of mother.
`weeks`          | length of pregnancy in weeks.
`premie`         | whether the birth was classified as premature (premie) or full-term.
`visits`         | number of hospital visits during pregnancy.
`marital`        | whether mother is `married` or `not married` at birth.
`gained`         | weight gained by mother during pregnancy in pounds.
`weight`         | weight of the baby at birth in pounds.
`lowbirthweight` | whether baby was classified as low birthweight (`low`) or not (`not low`).
`gender`         | gender of the baby, `female` or `male`.
`habit`          | status of the mother as a `nonsmoker` or a `smoker`.
`whitemom`       | whether mom is `white` or `not white`.

1.  What are the cases in this data set? How many cases are there in our sample?

Answer - The cases are for all the births population in North Carolina in 2004. Each case has the relevant information about the parents of the child, and some child's information as well.
The sample has 1000 cases.

As a first step in the analysis, we should consider summaries of the data. This
can be done using the `summary` command:

```{r summary, eval=TRUE}
summary(nc)
```

As you review the variable summaries, consider which variables are categorical 
and which are numerical. For numerical variables, are there outliers? If you 
aren't sure or want to take a closer look at the data, make a graph.

Answer:
Categorical variables are:
mature
premie
marital
lowbirthweight
gender
habit
whitemom

Numerical variables are:
fage
mage
weeks
visits
gained
weight

So out of total 13 variables, 7 are categorical and 6 are numerical.

Checking for the distribution of the numeric variables, and any outliers:
We are creating histograms for all the numeric variables below.

```{r}
hist(nc$fage)
hist(nc$mage)
hist(nc$weeks)
hist(nc$visits)
hist(nc$gained)
hist(nc$weight)
```

From the histogram plots, we will deduce the following:
1) Father's age - no outliers as per the histogram.
2) Mother's age - no outliers as per the histogram.
3) Lengths of pregeneancy is highly shewed on the lefyt. That clearly shows that there are outliers in this case.
4) Number of hospital visits is having outliers on the right, that means more number of visits.
5) Weight gained also has a strong right skew, and there are outliers on the right side.
6) Weight of the baby has strong skew on the left. It means there are some observations where new born babies have very less weights.

Consider the possible relationship between a mother's smoking habit and the 
weight of her baby. Plotting the data is a useful first step because it helps 
us quickly visualize trends, identify strong associations, and develop research
questions.

2.  Make a side-by-side boxplot of `habit` and `weight`. What does the plot 
highlight about the relationship between these two variables?

```{r}
qplot(nc$habit, nc$weight, geom = "boxplot", na.rm = TRUE)
```
The median weight of the new brons from the mothers who smoke is less than the same fromt eh mother who does not smoke. Even though there are many outliers on the lower side in case of the non-smoking mothers, but the general range of weight for the non-smoker mothers is slightly more than the same for the smoker mothers.

The box plots show how the medians of the two distributions compare, but we can
also compare the means of the distributions using the following function to 
split the `weight` variable into the `habit` groups, then take the mean of each
using the `mean` function.

```{r by-means, eval=TRUE}
by(nc$weight, nc$habit, mean)
```

There is an observed difference, but is this difference statistically 
significant? In order to answer this question we will conduct a hypothesis test
.

## Inference

3.  Check if the conditions necessary for inference are satisfied. Note that 
you will need to obtain sample sizes to check the conditions. You can compute 
the group size using the same `by` command above but replacing `mean` with 
`length`.

Answer: So, here we are dealing with the weight of the new borns under the 2 categories - smoking mothers and non-smoking mothers.
We will reate the histograms of the 2 categories separately:
```{r}
nc_smoking_mothers <- nc[nc$habit == "nonsmoker",]
nrow(nc_smoking_mothers)
hist(nc_smoking_mothers$weight)

nc_non_smoking_mothers <- nc[nc$habit == "smoker",]
nrow(nc_non_smoking_mothers)
hist(nc_non_smoking_mothers$weight)
```

To confirm the conditions necessary for inference:
1) Both the data sets - smomking and non-smking mothers are individually independent 
2) The distributions are skewed for both but the counts are large enough to ignore the skewness, and assume that this follows normal distribution.

4.  Write the hypotheses for testing if the average weights of babies born to 
smoking and non-smoking mothers are different.

Answer: 
Null Hypothesis H0        :  The difference between the average weight of babies born to smoking and no-smoking mothers is Zero
                              ??-non-smoking  -    ??-smoking    =    0
                              
Alternate  Hypothesis H0  :  The difference between the average weight of babies born to smoking and no-smoking mothers is not Zero
                              ??-non-smoking  -    ??-smoking    <>    0


Next, we introduce a new function, `inference`, that we will use for conducting
hypothesis tests and constructing confidence intervals. 

```{r inf-weight-habit-ht, eval=TRUE, tidy=FALSE}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```

Let's pause for a moment to go through the arguments of this custom function. 
The first argument is `y`, which is the response variable that we are 
interested in: `nc$weight`. The second argument is the explanatory variable, 
`x`, which is the variable that splits the data into two groups, smokers and 
non-smokers: `nc$habit`. The third argument, `est`, is the parameter we're 
interested in: `"mean"` (other options are `"median"`, or `"proportion"`.) Next
we decide on the `type` of inference we want: a hypothesis test (`"ht"`) or a 
confidence interval (`"ci"`). When performing a hypothesis test, we also need 
to supply the `null` value, which in this case is `0`, since the null 
hypothesis sets the two population means equal to each other. The `alternative`
hypothesis can be `"less"`, `"greater"`, or `"twosided"`. Lastly, the `method` 
of inference can be `"theoretical"` or `"simulation"` based.

5.  Change the `type` argument to `"ci"` to construct and record a confidence 
interval for the difference between the weights of babies born to smoking and 
non-smoking mothers.

By default the function reports an interval for ($\mu_{nonsmoker} - \mu_{smoker}$)
. We can easily change this order by using the `order` argument:

```{r inf-weight-habit-ci, eval=TRUE, tidy=FALSE}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical", 
          order = c("smoker","nonsmoker"))
```

* * *

## On your own

-   Calculate a 95% confidence interval for the average length of pregnancies 
(`weeks`) and interpret it in context. Note that since you're doing inference 
on a single population parameter, there is no explanatory variable, so you can 
omit the `x` variable from the function.

```{r}
inference(y = nc$weeks, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical")
```

Hence 95 % confidence level is: ( 38.1528 , 38.5165 )
That means that 95% of the women will have the average length falling in this range.


-   Calculate a new confidence interval for the same parameter at the 90% 
confidence level. You can change the confidence level by adding a new argument 
to the function: `conflevel = 0.90`.

```{r}
inference(y = nc$weeks, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical", conflevel = 0.90)
```


-   Conduct a hypothesis test evaluating whether the average weight gained by 
younger mothers is different than the average weight gained by mature mothers.

H0: average weight gained by younger mothers and average weight gained by mature mothers is same
    ??-weight-gained-younger-mothers - ??-weight-gained-mature-mothers = 0
H0: average weight gained by younger mothers and average weight gained by mature mothers is not same
    ??-weight-gained-younger-mothers - ??-weight-gained-mature-mothers <> 0

Now we have to find the p-value of the point estimate.

```{r}
xbar.weight.gained.younger.mothers <- mean(nc$gained[nc$mature == "younger mom"], na.rm = TRUE)

xbar.weight.gained.mature.mothers <- mean(nc$gained[nc$mature == "mature mom"], na.rm = TRUE)
point.mean.difference <- xbar.weight.gained.younger.mothers - xbar.weight.gained.mature.mothers
mean.null <- 0
s_young <- sd(nc$gained[nc$mature == "younger mom"], na.rm = TRUE)
n_young <- nrow(nc[nc$mature == "younger mom",])
s_mature <- sd(nc$gained[nc$mature == "mature mom"], na.rm = TRUE)
n_mature <- nrow(nc[nc$mature == "mature mom",])

SE.young.minus.mature <- sqrt(s_young ^2/n_young + s_mature ^2/n_mature)
z.value <- (point.mean.difference - mean.null) / SE.young.minus.mature
p.value <- 2 * pnorm(-abs(z.value))

```

Now since p-value > significance value which is normally taken as 0.05, H0 (null hypothesis) cannot be rejected.
Hence there is not much evidence to reject that average weight gained by younger mothers and average weight gained by mature mothers is same.


-   Now, a non-inference task: Determine the age cutoff for younger and mature 
mothers. Use a method of your choice, and explain how your method works.

```{r}
younger.mom.min.age <- nc$mage[nc$mature == "younger mom"] %>% min()
younger.mom.max.age <- nc$mage[nc$mature == "younger mom"] %>% max()
sprintf("Younger moms min age is: %s", younger.mom.min.age)
sprintf("Younger moms max age is: %s", younger.mom.max.age)

mature.mom.min.age <- nc$mage[nc$mature == "mature mom"] %>% min()
mature.mom.max.age <- nc$mage[nc$mature == "mature mom"] %>% max()
sprintf("Mature moms min age is: %s", mature.mom.min.age)
sprintf("Mature moms max age is: %s", mature.mom.max.age)
```


-   Pick a pair of numerical and categorical variables and come up with a 
research question evaluating the relationship between these variables. 
Formulate the question in a way that it can be answered using a hypothesis test
and/or a confidence interval. Answer your question using the `inference` 
function, report the statistical results, and also provide an explanation in 
plain language.

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0).
This lab was adapted for OpenIntro by Mine &Ccedil;etinkaya-Rundel from a lab 
written by the faculty and TAs of UCLA Statistics.
</div>
