---
title: "Multiple linear regression"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

## Grading the professor

Many college courses conclude by giving students the opportunity to evaluate 
the course and the instructor anonymously. However, the use of these student 
evaluations as an indicator of course quality and teaching effectiveness is 
often criticized because these measures may reflect the influence of 
non-teaching related characteristics, such as the physical appearance of the 
instructor. The article titled, "Beauty in the classroom: instructors' 
pulchritude and putative pedagogical productivity" (Hamermesh and Parker, 2005) 
found that instructors who are viewed to be better looking receive higher 
instructional ratings. (Daniel S. Hamermesh, Amy Parker, Beauty in the 
classroom: instructors pulchritude and  putative pedagogical productivity, 
*Economics of Education Review*, Volume 24, Issue 4, August 2005, Pages 369-376, 
ISSN 0272-7757, 10.1016/j.econedurev.2004.07.013. [http://www.sciencedirect.com/science/article/pii/S0272775704001165](http://www.sciencedirect.com/science/article/pii/S0272775704001165).)

In this lab we will analyze the data from this study in order to learn what goes 
into a positive professor evaluation.

## The data

The data were gathered from end of semester student evaluations for a large 
sample of professors from the University of Texas at Austin. In addition, six 
students rated the professors' physical appearance. (This is aslightly modified 
version of the original data set that was released as part of the replication 
data for *Data Analysis Using Regression and Multilevel/Hierarchical Models* 
(Gelman and Hill, 2007).) The result is a data frame where each row contains a 
different course and columns represent variables about the courses and professors.

```{r load-data, eval=TRUE}
rm(list = ls())
load("more/evals.RData")
```

variable         | description
---------------- | -----------
`score`          | average professor evaluation score: (1) very unsatisfactory - (5) excellent.
`rank`           | rank of professor: teaching, tenure track, tenured.
`ethnicity`      | ethnicity of professor: not minority, minority.
`gender`         | gender of professor: female, male.
`language`       | language of school where professor received education: english or non-english.
`age`            | age of professor.
`cls_perc_eval`  | percent of students in class who completed evaluation.
`cls_did_eval`   | number of students in class who completed evaluation.
`cls_students`   | total number of students in class.
`cls_level`      | class level: lower, upper.
`cls_profs`      | number of professors teaching sections in course in sample: single, multiple.
`cls_credits`    | number of credits of class: one credit (lab, PE, etc.), multi credit.
`bty_f1lower`    | beauty rating of professor from lower level female: (1) lowest - (10) highest.
`bty_f1upper`    | beauty rating of professor from upper level female: (1) lowest - (10) highest.
`bty_f2upper`    | beauty rating of professor from second upper level female: (1) lowest - (10) highest.
`bty_m1lower`    | beauty rating of professor from lower level male: (1) lowest - (10) highest.
`bty_m1upper`    | beauty rating of professor from upper level male: (1) lowest - (10) highest.
`bty_m2upper`    | beauty rating of professor from second upper level male: (1) lowest - (10) highest.
`bty_avg`        | average beauty rating of professor.
`pic_outfit`     | outfit of professor in picture: not formal, formal.
`pic_color`      | color of professor's picture: color, black & white.

## Exploring the data

1.  Is this an observational study or an experiment? The original research
    question posed in the paper is whether beauty leads directly to the
    differences in course evaluations. Given the study design, is it possible to
    answer this question as it is phrased? If not, rephrase the question.

Answer: This is an observational study, as this is already a process which is followed, and t was not specially done for this analysis.

2.  Describe the distribution of `score`. Is the distribution skewed? What does 
    that tell you about how students rate courses? Is this what you expected to 
    see? Why, or why not?
Answer: 
```{r}
hist(evals$score)
```

As we see from the histogram, the score is strongly skewed on the left. That means students prefer not to rate the professors below a certain level. Also we see that most of the professors are rated towards the highest rating.

3.  Excluding `score`, select two other variables and describe their relationship 
    using an appropriate visualization (scatterplot, side-by-side boxplots, or 
    mosaic plot).

Answer: We will use 2 variables: cls_perc_eval and bty_avg. Both of these are continuous numeric variables. And we can try to see if the average beauty rating of a professor and the percentage of students in the class who voted are related to each other.
Creating a scatter plot to see how the relationship looks like:
```{r}
library(ggplot2)

ggplot(evals, aes(x=bty_avg, y=cls_perc_eval)) +
  geom_point()

bty_avg.cls_perc_eval.model <- lm(formula = cls_perc_eval ~ bty_avg, data = evals)
summary(bty_avg.cls_perc_eval.model)
```

From above, it is clear that there is statistical relationship between the 2 variables we chose above, but the R value is quite less.

## Simple linear regression

The fundamental phenomenon suggested by the study is that better looking teachers
are evaluated more favorably. Let's create a scatterplot to see if this appears 
to be the case:

```{r scatter-score-bty_avg, eval = FALSE}
plot(evals$score ~ evals$bty_avg)
```


Before we draw conclusions about the trend, compare the number of observations 
in the data frame with the approximate number of points on the scatterplot. 
Is anything awry?
Answer: Apparently, as the average beauty range increases, the score is also raising.
The number of points on the plot does not look to be equal to the number of observations which is 463. Some points seem to overlap hence looking like a lesser number.

4.  Replot the scatterplot, but this time use the function `jitter()` on the
    $y$- or the $x$-coordinate. (Use `?jitter` to learn more.) What was
    misleading about the initial scatterplot?
Answer:

```{r}
plot(evals$score ~ jitter(evals$bty_avg, 1))

```
This gives jitter (noise) on the x-axis, meaning the points are made apart so that more number of points could be plotted and are visible.

5.  Let's see if the apparent trend in the plot is something more than
    natural variation. Fit a linear model called `m_bty` to predict average
    professor score by average beauty rating and add the line to your plot
    using `abline(m_bty)`. Write out the equation for the linear model and
    interpret the slope. Is average beauty score a statistically significant
    predictor? Does it appear to be a practically significant predictor?
    
Answer:
```{r}
m_bty <- lm(formula = score ~ bty_avg, data = evals)

summary(m_bty)
```

Slope is a very less value : 0.06664. That means for increase of 1 score in average beauty, the score increases by 0.06664

```{r}
ggplot(evals, aes(x=bty_avg, y=score)) +
  geom_point(color = "red") +
  geom_line(aes(x=bty_avg, y=predict(m_bty, newdata = evals)), color = "blue")
```


6.  Use residual plots to evaluate whether the conditions of least squares
    regression are reasonable. Provide plots and comments for each one (see
    the Simple Regression Lab for a reminder of how to make these).

Plotting the residual plot:
```{r}
plot(m_bty$residuals ~ evals$bty_avg)
abline(h = 0, lty = 3)
```

Creating the residuals histogram
```{r}
hist(m_bty$residuals)
```

Q-Q Plot for residuals:
```{r}
qqnorm(m_bty$residuals)
qqline(m_bty$residuals)
```

Let us evaluate the 4 conditions for fitting the least squared line for this scenario:
1) Linearity: From the linear model and the plot made from the lm model, it is clear that the data trend is linear for this one. Hence this condition is satisfied.
2) Nearly normal residuals: As seen from the histogram, the residuals are not nearly normal for this. Hence this condition is not satisfied too.
3) Constant Variability: From the residuals plot drawn above, the variability s not constant as the positive residuals vary from 0 to 1 which the negative ones go as far as -2. Hence the constant variability condition is also not satisfied.
4) Independent observations: The observations can be said to be independent as this is not a time based observation.

As the conditions for normal residuals and constant variability are not satisfied, we can say that the least squared line cannot be fitted for this scenario.


## Multiple linear regression

The data set contains several variables on the beauty score of the professor: 
individual ratings from each of the six students who were asked to score the 
physical appearance of the professors and the average of these six scores. Let's 
take a look at the relationship between one of these scores and the average 
beauty score.

```{r bty-rel, eval = FALSE}
plot(evals$bty_avg ~ evals$bty_f1lower)
cor(evals$bty_avg, evals$bty_f1lower)
```

As expected the relationship is quite strong - after all, the average score is 
calculated using the individual scores. We can actually take a look at the 
relationships between all beauty variables (columns 13 through 19) using the 
following command:

```{r bty-rels, eval = FALSE}
plot(evals[,13:19])
```

These variables are collinear (correlated), and adding more than one of these 
variables to the model would not add much value to the model. In this 
application and with these highly-correlated predictors, it is reasonable to use
the average beauty score as the single representative of these variables.

In order to see if beauty is still a significant predictor of professor score 
after we've accounted for the gender of the professor, we can add the gender 
term into the model.

```{r scatter-score-bty_avg_gender, include=TRUE}
m_bty_gen <- lm(score ~ bty_avg + gender, data = evals)
summary(m_bty_gen)
```

7.  P-values and parameter estimates should only be trusted if the
    conditions for the regression are reasonable. Verify that the conditions
    for this model are reasonable using diagnostic plots.


Answer:

Building the residual plot:
```{r}
plot(m_bty_gen$residuals ~ evals$bty_avg)
abline(h = 0, lty = 3)
```

Histogram of residuals:
```{r}
hist(m_bty_gen$residuals)
```

Q-Q Plot:
```{r}
qqnorm(m_bty_gen$residuals)
qqline(m_bty_gen$residuals)
```

Checking for the regression:
1) Linearity: The data looks to be linear.
2)Nearly normal Residuals: From the residuals histogram, it is clear the histogram is skewed on the left, and is not nearly normal.
3) Constant Variability: From the residual plot, similar to the previous one with the model with only bty_avg, the variability is not constant.
4) Indenpendent Observations : The observations are all independent.

However 1 and 4 conditions are met, 2 and 4 are not satisfied. Hence all the conditions are not met.

8.  Is `bty_avg` still a significant predictor of `score`? Has the addition
    of `gender` to the model changed the parameter estimate for `bty_avg`?

Note that the estimate for `gender` is now called `gendermale`. You'll see this 
name change whenever you introduce a categorical variable. The reason is that R 
recodes `gender` from having the values of `female` and `male` to being an 
indicator variable called `gendermale` that takes a value of $0$ for females and
a value of $1$ for males. (Such variables are often referred to as "dummy" 
variables.)

As a result, for females, the parameter estimate is multiplied by zero, leaving 
the intercept and slope form familiar from simple regression.

\[
  \begin{aligned}
\widehat{score} &= \hat{\beta}_0 + \hat{\beta}_1 \times bty\_avg + \hat{\beta}_2 \times (0) \\
&= \hat{\beta}_0 + \hat{\beta}_1 \times bty\_avg\end{aligned}
\]

We can plot this line and the line corresponding to males with the following 
custom function.

```{r twoLines, eval = FALSE}
multiLines(m_bty_gen)
```

9.  What is the equation of the line corresponding to males? (*Hint:* For
    males, the parameter estimate is multiplied by 1.) For two professors
    who received the same beauty rating, which gender tends to have the
    higher course evaluation score?

The decision to call the indicator variable `gendermale` instead of`genderfemale`
has no deeper meaning. R simply codes the category that comes first 
alphabetically as a $0$. (You can change the reference level of a categorical 
variable, which is the level that is coded as a 0, using the`relevel` function. 
Use `?relevel` to learn more.)

Answer:

score = 3.75 + 0.074 X bty_avg + 0.173 X gender
where gender = 1 for male
and gender = 0 for female

For males:
score = 3.75 + 0.074 X bty_avg + 0.173


10. Create a new model called `m_bty_rank` with `gender` removed and `rank` 
    added in. How does R appear to handle categorical variables that have more 
    than two levels? Note that the rank variable has three levels: `teaching`, 
    `tenure track`, `tenured`.

```{r}
m_bty_rank <- lm(score ~ bty_avg + rank, data = evals)
summary(m_bty_rank)

multiLines(m_bty_rank)
```

R adds (n - 1) variables for a categorical value with n categories.

The interpretation of the coefficients in multiple regression is slightly 
different from that of simple regression. The estimate for `bty_avg` reflects
how much higher a group of professors is expected to score if they have a beauty
rating that is one point higher *while holding all other variables constant*. In
this case, that translates into considering only professors of the same rank 
with `bty_avg` scores that are one point apart.

## The search for the best model

We will start with a full model that predicts professor score based on rank, 
ethnicity, gender, language of the university where they got their degree, age, 
proportion of students that filled out evaluations, class size, course level, 
number of professors, number of credits, average beauty rating, outfit, and 
picture color.

11. Which variable would you expect to have the highest p-value in this model? 
    Why? *Hint:* Think about which variable would you expect to not have any 
    association with the professor score.

Answer: The variable pic_outfit is expected to be not a predictor for the score.

Let's run the model...

```{r m_full, eval = FALSE, tidy = FALSE}
m_full <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_profs + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m_full)
```

12. Check your suspicions from the previous exercise. Include the model output
    in your response.

Answer:
I was expecting the variable cls_profs (number of professors teaching sections in course in sample: single, multiple.) to be at least at some level a predictor of the score. But as we see the p-value of this variable is 0.77 which is very high. That means this variable cannot be a predictor of the score.

13. Interpret the coefficient associated with the ethnicity variable.
Answer: 
The coefficient for ethnicity of non-minority is 0.12 while 0 in case of a minority. That means for professors from non-minority have a score of 0.12 greater than the score for a minor  ethiicity background professor. 

14. Drop the variable with the highest p-value and re-fit the model. Did the
    coefficients and significance of the other explanatory variables change?
    (One of the things that makes multiple regression interesting is that
    coefficient estimates depend on the other variables that are included in
    the model.) If not, what does this say about whether or not the dropped
    variable was collinear with the other explanatory variables?

Answer:Highest p-value from the full model is for cls_profssingle. Hence removing the variable cls_profs, and rebuilding the model.
```{r}
m_full_minus_cls_profs <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
                             + cls_students + cls_level + cls_credits + bty_avg 
                             + pic_outfit + pic_color, data = evals)
summary(m_full_minus_cls_profs)
```


15. Using backward-selection and p-value as the selection criterion,
    determine the best model. You do not need to show all steps in your
    answer, just the output for the final model. Also, write out the linear
    model for predicting score based on the final model you settle on.

Answer:
Moving forward with the m_full_minus_cls_profs. The highest p-value is for the variable: cls_levelupper which is 0.29. Hence removing this variable cls_level

```{r}
model2 <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(model2)
```

Now we see that the highest p-value is for cls_students (0.31) which is greater than SL of 0.05. Hence moving forward and removing this field from the regression model:
```{r}
model3 <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
             + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(model3)
```

Now we see that the variable ranktenuredhas the highest p-value of 0.2 which is greater than 0.05. Hence removing this column:
```{r}
model4 <- lm(score ~ ethnicity + gender + language + age + cls_perc_eval 
             + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(model4)
```

Now wee that the column pic_outfitnot formal has the highest p-value of 0.09 which is greater than 0.05. Hence removing this field:
```{r}
model5 <- lm(score ~ ethnicity + gender + language + age + cls_perc_eval 
             + cls_credits + bty_avg 
             + pic_color, data = evals)
summary(model5)
```

Now the biggest p-value is for languagenon-english which is 0.047. As it is less than 0.05, it can be ignored. And this is the final model.

16. Verify that the conditions for this model are reasonable using diagnostic 
    plots.

Answer:
Building the residual plot:
```{r}
plot(model5$residuals ~ model5$fitted.values)
abline(h = 0, lty = 3)
```

Histogram of residuals:
```{r}
hist(model5$residuals)
```

Q-Q Plot:
```{r}
qqnorm(model5$residuals)
qqline(model5$residuals)
```

Checking for the regression:
1) Linearity: The data looks to be linear.
2)Nearly normal Residuals: From the residuals histogram, it is clear the histogram is skewed on the left, and is not nearly normal.
3) Constant Variability: From the residual plot, similar to the previous one with the model with only bty_avg, the variability is not constant.
4) Indenpendent Observations : The observations are all independent.


17. The original paper describes how these data were gathered by taking a
    sample of professors from the University of Texas at Austin and including 
    all courses that they have taught. Considering that each row represents a 
    course, could this new information have an impact on any of the conditions 
    of linear regression?
Answer: It might be impacted. There might be multiple courses taken by the same professor, so in that case the sameprofessor might have got 2 different scores for the 2 different courses he or she takes.

18. Based on your final model, describe the characteristics of a professor and 
    course at University of Texas at Austin that would be associated with a high
    evaluation score.

Answer:
High evaluation score means the categorical variables with the positive values have the category as represented by the model, and the categorical variables with a negative coefficient has the other category, so that the score becomes maximum.
Hence, ethnicity of non-minority, gender male, language english, cls_credit of one level, pic_color value as black & white - will be the ones with the highest scores.

19. Would you be comfortable generalizing your conclusions to apply to professors
    generally (at any university)? Why or why not?

Answer: No, the conclusions cannot be generalized to other universities, as there can be many other factors impacting the scores.

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0). This lab was written by 
Mine &Ccedil;etinkaya-Rundel and Andrew Bray.
</div>
